---
title: "aggression_vs_scoring"
output: html_document
---
approaches based on aggression...
do more aggressive players also score more points?
scatterplot of points (sum of goals and assists) vs. cards (sum of red and yellow)

because defenders tend not to score frequently, the defenders are filtered out.

(I can't decide between geom_point(), which looks cleaner, and geom_jitter(), which gives more of a sense of the number of observations...)

geom_jitter() version
```{r}
nwsl %>%
  filter(pos %in% c("MF", "FW,MF","FW","DF,MF","DF,FW")) %>%
  mutate(points = goals + assists, cards = yellow_cards + red_cards) %>%
  ggplot(aes(x = cards, y = points, color = pos)) +
  geom_jitter() + 
  stat_smooth(method = "lm")+
  theme_bw() +
  labs(title = "Aggression and Scoring", x = "Red and Yellow Cards", y = "Points")
```
geom_point() version
```{r}
nwsl %>%
  filter(pos %in% c("MF", "FW,MF","FW","DF,MF","DF,FW")) %>%
  mutate(points = goals + assists, cards = yellow_cards + red_cards) %>%
  ggplot(aes(x = cards, y = points, color = pos)) +
  geom_point() +
  stat_smooth(method = "lm")+
  theme_bw() +
  labs(title = "Aggression and Scoring", x = "Red and Yellow Cards", y = "Points")
```

