---
title: "aggression_vs_scoring"
output: html_document
---
approaches based on aggression...
do more aggressive players also score more points?
scatterplot of points (sum of goals and assists) vs. cards (sum of red and yellow)

because defenders tend not to score frequently, the defenders are filtered out.

(I can't decide between geom_point(), which looks cleaner, and geom_jitter(), which gives more of a sense of the number of observations...)

geom_jitter() version
```{r}
aggression_scoring_plot <- nwsl %>%
  filter(pos %in% c("MF", "FW,MF","FW","DF,MF","DF,FW")) %>%
  mutate(points = goals + assists, cards = yellow_cards + red_cards) %>%
  ggplot(aes(x = cards, y = points, color = pos)) +
  geom_jitter() + 
  stat_smooth(method = "lm")+
  theme_bw() +
  labs(title = "Greater Aggression Relates to Greater Scoring in Forwards", subtitle = "2017-2019 Regular Season Data, Excluding Defenders", x = "Red and Yellow Cards", y = "Points", color = "Position")

aggression_scoring_plot

ggsave("../Images/aggression_scoring_plot.jpg", width = 8, height = 6)
```
geom_point() version
```{r}
nwsl %>%
  filter(pos %in% c("MF", "FW,MF","FW","DF,MF","DF,FW")) %>%
  mutate(points = goals + assists, cards = yellow_cards + red_cards) %>%
  ggplot(aes(x = cards, y = points, color = pos)) +
  geom_point() +
  stat_smooth(method = "lm")+
  theme_bw() +
  labs(title = "Aggression and Scoring", x = "Red and Yellow Cards", y = "Points")
```

