---
title: "exploration2"
output: html_document
---


An R markdown file for exploring the NWSL data!

```{r}
library(tidyverse)
nwsl <- read.csv("../Data/nwsl_season_stats.csv")
str(nwsl)
```

Generate some exploratory graphs...

looking at goals by position
```{r}
nwsl %>%
  ggplot(aes(x = goals, y = pos)) +
  geom_violin()
```
how many players are there at each position anyway?

```{r}
nwsl %>%
  group_by(pos) %>%
  summarize(count = n())
```
maybe it would be better to make the people who are two positions actually be only one position? to clean the data?


approaches based on aggression...
do more aggressive players also score more points?
scatterplot of points (sum of goals and assists) vs. cards (sum of red and yellow)

```{r}
nwsl %>%
  mutate(points = goals + assists, cards = yellow_cards + red_cards) %>%
  ggplot(aes(x = cards, y = points, color = pos)) +
  geom_jitter() +
  theme_bw() +
  labs(title = "Aggression and Scoring", x = "Red and Yellow Cards", y = "Points")
```

filtering out the defenders (and the 12 "DF,FW") as well as all of the players who did not get either a point or an assist over the season


```{r}
nwsl %>%
  filter(pos %in% c("MF", "FW,MF","FW","DF,MF","DF,FW"), ((goals>0)|(assists>0))) %>%
  mutate(points = goals + assists, cards = yellow_cards + red_cards) %>%
  ggplot(aes(x = cards, y = points, color = pos)) +
  geom_jitter() +
  stat_smooth(method = "lm")+
  theme_bw() +
  labs(title = "Aggression and Scoring", x = "Red and Yellow Cards", y = "Points")
```
but would I rather leave in the players who never scored?

```{r}
nwsl %>%
  filter(pos %in% c("MF", "FW,MF","FW","DF,MF","DF,FW")) %>%
  mutate(points = goals + assists, cards = yellow_cards + red_cards) %>%
  ggplot(aes(x = cards, y = points, color = pos)) +
  geom_jitter() + 
  stat_smooth(method = "lm")+
  theme_bw() +
  labs(title = "Aggression and Scoring", x = "Red and Yellow Cards", y = "Points")
```
or is geom_point clearer??

```{r}
nwsl %>%
  filter(pos %in% c("MF", "FW,MF","FW","DF,MF","DF,FW")) %>%
  mutate(points = goals + assists, cards = yellow_cards + red_cards) %>%
  ggplot(aes(x = cards, y = points, color = pos)) +
  geom_point() +
  stat_smooth(method = "lm")+
  theme_bw() +
  labs(title = "Aggression and Scoring", x = "Red and Yellow Cards", y = "Points")
```

do different teams have different approaches with regards to star players versus lots of low-scoring players?

first off, what does the histogram of goals per player look like for one team, say BOS?

```{r}
nwsl %>%
  filter(team_id == "BOS") %>%
  ggplot(aes(x = goals)) +
  geom_histogram()
```
maybe I should filter by the amount of time that they've played, because the number that scored zero is so high?

```{r}
nwsl %>%
  filter(team_id == "BOS", minutes >= 100) %>%
  ggplot(aes(x = goals)) +
  geom_histogram()
```


```{r}
nwsl %>%
  ggplot(aes(x = goals, y = team_id)) +
  geom_violin() +
  geom_boxplot(width = 0.2) +
  theme_bw()
```
back to approaches based on aggression...
can we break the players into groups based on style of play with regards to aggression and scoring?

```{r}
nwsl <- nwsl %>%
  mutate(points = goals + assists, cards = red_cards + yellow_cards)

soccer_clustering <- protoclust(dist(dplyr::select(nwsl, 
                                                   points, cards)))
plot(soccer_clustering)
  
```
well that dendrogram is not helpful...

```{r}
minimax_player_clusters <- protocut(soccer_clustering, k = 3)

nwsl %>%
  mutate(player_style_clusters = as.factor(minimax_player_clusters$cl)) %>%
  ggplot(aes(x = cards, y = points, color = player_style_clusters)) +
  geom_point() +
  theme_bw()
```

so it looks like clustering ends up being really about point-scoring and not really about getting red or yellow cards

what if I have it do more clusters?

```{r}
minimax_player_clusters <- protocut(soccer_clustering, k = 4)

nwsl %>%
  mutate(player_style_clusters = as.factor(minimax_player_clusters$cl)) %>%
  ggplot(aes(x = cards, y = points, color = player_style_clusters)) +
  geom_point() +
  theme_bw()
```

four looks pretty much the same...

```{r}
minimax_player_clusters <- protocut(soccer_clustering, k = 5)

nwsl %>%
  mutate(player_style_clusters = as.factor(minimax_player_clusters$cl)) %>%
  ggplot(aes(x = cards, y = points, color = player_style_clusters)) +
  geom_point() +
  theme_bw()
```
I think I would stick with the 3 cluster example, just because there are three main positions in soccer (D, M, F), but it might be good to add shapes of the points based on positions...

```{r}
minimax_player_clusters <- protocut(soccer_clustering, k = 3)

nwsl %>%
  mutate(player_style_clusters = as.factor(minimax_player_clusters$cl)) %>%
  ggplot(aes(x = cards, y = points, color = player_style_clusters,
             shape = pos)) +
  geom_jitter() +
  theme_bw()
```
I think I will go back and make these all restricted to players who were above a certain percentile of playing time...

```{r}
nwsl %>%
  ggplot(aes(x = minutes)) +
  stat_ecdf() +
  theme_bw()
```
the ecdf is not really showing anything jumping out at me... so I guess I'll just go with 100 minutes and see if that has an effect...


```{r}
main_players_nwsl <- nwsl %>%
  filter(minutes >= 100)

soccer_clustering <- protoclust(dist(dplyr::select(main_players_nwsl, 
                                                   points, cards)))

minimax_player_clusters <- protocut(soccer_clustering, k = 3)

main_players_nwsl %>%
  mutate(player_style_clusters = as.factor(minimax_player_clusters$cl)) %>%
  ggplot(aes(x = cards, y = points, color = player_style_clusters,
             shape = pos)) +
  geom_jitter() +
  theme_bw()

```
but there is still a lot going on in that bottom left corner, so maybe I should increase my threshold for the number of minutes they have to play...

```{r}
main_players_nwsl <- nwsl %>%
  filter(minutes >= 150)

soccer_clustering <- protoclust(dist(dplyr::select(main_players_nwsl, 
                                                   points, cards)))

minimax_player_clusters <- protocut(soccer_clustering, k = 3)

main_players_nwsl %>%
  mutate(player_style_clusters = as.factor(minimax_player_clusters$cl)) %>%
  ggplot(aes(x = cards, y = points, color = player_style_clusters,
             shape = pos)) +
  geom_jitter() +
  theme_bw()
```
that is a bit better, but what if I go up even more...

```{r}
main_players_nwsl <- nwsl %>%
  filter(minutes >= 200)

soccer_clustering <- protoclust(dist(dplyr::select(main_players_nwsl, 
                                                   points, cards)))

minimax_player_clusters <- protocut(soccer_clustering, k = 3)

main_players_nwsl %>%
  mutate(player_style_clusters = as.factor(minimax_player_clusters$cl)) %>%
  ggplot(aes(x = cards, y = points, color = player_style_clusters,
             shape = pos)) +
  geom_jitter() +
  theme_bw()
```
I think I'll go with that one...

