---
title: "player_style_clustering"
output: html_document
---
again looking at aggression and scoring, can we break the players into groups based on style of play with regards to aggression and scoring?

I tried clustering based on points and cards with 3, 4, and 5 clusters, and they all looked relatively similar, showing that clustering ends up being really about point-scoring and not really about getting red or yellow cards. I decided to stick with the three-cluster example because there are three main (non-goalie) positions in soccer.

In the scatterplot, I also incorporated shape of the points to represent the actual position the players were to see if there was an association between the clustering and actual positions.

Finally, I also restricted the players considered to those who played more than 200 minutes to eliminate some of the noise in the bottom left corner of players who didn't really play enough to have an impact.

```{r}
main_players_nwsl <- nwsl %>%
  filter(minutes >= 200)

soccer_clustering <- protoclust(dist(dplyr::select(main_players_nwsl, 
                                                   points, cards)))

minimax_player_clusters <- protocut(soccer_clustering, k = 3)

main_players_nwsl %>%
  mutate(player_style_clusters = as.factor(minimax_player_clusters$cl)) %>%
  ggplot(aes(x = cards, y = points, color = player_style_clusters,
             shape = pos)) +
  geom_jitter() +
  labs(title = "Clustering of Player Styles", x = "Red and Yellow Cards", 
       y = "Points", shape = "Position", color = "Player Style Cluster") +
  theme_bw()
```
now, to see which players are the prototypes for each cluster...

```{r}
minimax_player_clusters$protos

dplyr::select(main_players_nwsl[minimax_player_clusters$protos, ], name, points, cards)
```
So the prototypical player for cluster 1 is Casey Short (a defender), cluster 2 is Christine Sinclair (a FW,MF), and cluster 3 is Sam Kerr (also a FW,MF).
